---
title: "DDSAnalytics Talent Management EDA "
output: html_document
editor_options: 
  chunk_output_type: inline
---
## Web link to Youtube video
<a href="https://youtu.be/Y9-0hh8LNzg"> Cleveland's Youtube </a>  
  
## Introduction

Today we will use the our data set to answer the following question:  
* What does our data look like?  
* What data was removed?  
* What factors lead to attrition? What are the top 3?  
* Can we build a model to predict attrition?  
* What factors contribute to monthly income?  
* Can we build a model to predict monthly income?  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

##Initial Data Check  
We found 870 observations with 36 variables.  
19 Continous variables.  
9 Nominal variables.  
8 Ordinal variables.  
The 8 Ordinal variables were imported ad intergers and were converted to Factors.
  
Standard Hours, Over18, and EmployeeCount were not considered during EDA.  
These variables had the same value for all observations.  
```{r Bring in Libraries + Load the data}
library(tidyverse)
library(readxl)
library(curl)
library(GGally)
library(e1071)
library(caret)
library(ggcorrplot)
library(gridExtra)

data <-read.csv(curl("https://raw.githubusercontent.com/tricleve/CaseStudy2DDS/master/CaseStudy2-data.csv"))

#str(data)
#summary(data)
#No missing values were found

#Ordinal data was captured as integers.  Need to change variable type to Factor
data$Education = as.factor(data$Education)
data$EnvironmentSatisfaction = as.factor(data$EnvironmentSatisfaction)
data$JobInvolvement = as.factor(data$JobInvolvement)
data$JobLevel = as.factor(data$JobLevel)
data$JobSatisfaction = as.factor(data$JobSatisfaction)
data$PerformanceRating = as.factor(data$PerformanceRating)
data$RelationshipSatisfaction = as.factor(data$RelationshipSatisfaction)
data$WorkLifeBalance = as.factor(data$WorkLifeBalance)


#Import Attrition Competition set and change variables
Comp_attr <- read.csv(curl("https://raw.githubusercontent.com/tricleve/CaseStudy2DDS/master/CaseStudy2CompSetNoAttrition.csv"))
Comp_attr$Education = as.factor(Comp_attr$Education)
Comp_attr$EnvironmentSatisfaction = as.factor(Comp_attr$EnvironmentSatisfaction)
Comp_attr$JobInvolvement = as.factor(Comp_attr$JobInvolvement)
Comp_attr$JobLevel = as.factor(Comp_attr$JobLevel)
Comp_attr$JobSatisfaction = as.factor(Comp_attr$JobSatisfaction)
Comp_attr$PerformanceRating = as.factor(Comp_attr$PerformanceRating)
Comp_attr$RelationshipSatisfaction = as.factor(Comp_attr$RelationshipSatisfaction)
Comp_attr$WorkLifeBalance = as.factor(Comp_attr$WorkLifeBalance)
#str(Comp_attr)

#Import Salary Competition data set adn change variables
Comp_sal <-read.csv(curl("https://raw.githubusercontent.com/tricleve/CaseStudy2DDS/master/CaseStudy2CompSetNoSalary.csv"))
Comp_sal$Education = as.factor(Comp_sal$Education)
Comp_sal$EnvironmentSatisfaction = as.factor(Comp_sal$EnvironmentSatisfaction)
Comp_sal$JobInvolvement = as.factor(Comp_sal$JobInvolvement)
Comp_sal$JobLevel = as.factor(Comp_sal$JobLevel)
Comp_sal$JobSatisfaction = as.factor(Comp_sal$JobSatisfaction)
Comp_sal$PerformanceRating = as.factor(Comp_sal$PerformanceRating)
Comp_sal$RelationshipSatisfaction = as.factor(Comp_sal$RelationshipSatisfaction)
Comp_sal$WorkLifeBalance = as.factor(Comp_sal$WorkLifeBalance)
#str(Comp_sal)

```

## Attrition: Continuous EDA
```{r Attrition: Continous EDA}
#EDA of Continuous variables in data set

#Appears to be a factor
MI_plot <-data %>% ggplot(aes(MonthlyIncome))+geom_boxplot(aes(fill=Attrition))+labs(title="Monthly Income vs Attrition")
TWY_plot <-data %>% ggplot(aes(TotalWorkingYears))+geom_boxplot(aes(fill=Attrition))+labs(title="TotalWorkingYears vs Attrition")
YAC_plot <- data %>% ggplot(aes(YearsAtCompany))+geom_boxplot(aes(fill=Attrition))+labs(title="YearsAtCompany vs Attrition")
YICR_plot <- data %>% ggplot(aes(YearsInCurrentRole))+geom_boxplot(aes(fill=Attrition))+labs(title="YearsInCurrentRole vs Attrition")
YWCM_plot <- data %>% ggplot(aes(YearsWithCurrManager))+geom_boxplot(aes(fill=Attrition))+labs(title="YearsWithCurrManager vs Attrition")

#T-tests for continuous variables that appear to be significant(ranked in ascending P-value order)
leave_MI <- data %>% filter(Attrition=="Yes") %>% select(MonthlyIncome)
stay_MI <- data %>% filter(Attrition=="No") %>% select(MonthlyIncome)
MI_t <-t.test(leave_MI, stay_MI, alternative="two.sided")

leave_TWY <- data %>% filter(Attrition=="Yes") %>% select(TotalWorkingYears)
stay_TWY <- data %>% filter(Attrition=="No") %>% select(TotalWorkingYears)
TWY_t <-t.test(leave_TWY, stay_TWY, alternative="two.sided")


leave_YICR <- data %>% filter(Attrition=="Yes") %>% select(YearsInCurrentRole)
stay_YICR <- data %>% filter(Attrition=="No") %>% select(YearsInCurrentRole)
YICR <- t.test(leave_YICR, stay_YICR, alternative="two.sided")

leave_YWCM <- data %>% filter(Attrition=="Yes") %>% select(YearsWithCurrManager)
stay_YWCM <- data %>% filter(Attrition=="No") %>% select(YearsWithCurrManager)
YWCM <- t.test(leave_YWCM, stay_YWCM, alternative="two.sided")


leave_YAC <- data %>% filter(Attrition=="Yes") %>% select(YearsAtCompany)
stay_YAC <- data %>% filter(Attrition=="No") %>% select(YearsAtCompany)
YAC <- t.test(leave_YAC, stay_YAC, alternative="two.sided")

Cont_var = c("MonthlyIncome", "TotalWorkingYears","YearsInCurrentRole","YearsWithCurrManager","YearsAtCompany")
ttest_p = c(MI_t$p.value, TWY_t$p.value, YICR$p.value, YWCM$p.value, YAC$p.value)

df_ttest = data.frame(Variable=Cont_var, "TTest pvalue"=ttest_p)
```
Boxplots were used to determine a relationship between the continuonus variables and Attrition.  
Five variables appears signficant:  
* Monthly Income  
* TotalWOrkingYears  
* YearsAtCompany  
* YearsInCurrentRold  
* YearsWithCurrManager  

A t-test was conducted to confirm signficance for these 5 variables.  
The boxplots and p value table is below.  
```{r Attrition: Continuous EDA cont.}
grid.arrange(MI_plot, TWY_plot, YAC_plot,YICR_plot, YWCM_plot,tableGrob(df_ttest))

#Doesn't appear to be a factor
#data %>% ggplot(aes(Age))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(YearsSinceLastPromotion))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(TrainingTimesLastYear))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(StockOptionLevel))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(PercentSalaryHike))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(NumCompaniesWorked))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(MonthlyRate))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(HourlyRate))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(DailyRate))+geom_boxplot(aes(fill=Attrition))
#data %>% ggplot(aes(DistanceFromHome))+geom_boxplot(aes(fill=Attrition))
```
  
## Attrition: Categorical EDA  
The relationship between categorical variables and Attrition was assessed using a Proportion tables.
A sample table is below:  

```{r Attrition: Categorical EDA}
#Categorical variable EDA

#Categorical variable analysis -> Looking at percentage left by each response
#Appears to be a factor
data %>% group_by(JobLevel) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(OverTime) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(JobInvolvement) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(JobRole) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(MaritalStatus) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")

```
  
5 Categorical variables were identified as having a relationship with Attrition:  
* JobLevel  
* OverTime  
* JobInvolvement  
* JobSatisfaction  
* JobRole  
*  Marital Status  
A bar chart vs Attrition and a chi square test was also completed for each of these variables.  
The results can be seen below:
```{r Attrition: Categorical EDA cont}
#Bar plots for variables deemed factors
JL_plot <- data %>% ggplot(aes(JobLevel))+geom_bar(aes(fill=Attrition)) + labs(title="JobLevel vs Attrition")
O_plot <- data %>% ggplot(aes(OverTime))+geom_bar(aes(fill=Attrition)) + labs(title="Overtime vs Attrition")
JI_plot <- data %>% ggplot(aes(JobInvolvement))+geom_bar(aes(fill=Attrition)) + labs(title="JobInvolvement vs Attrition")
JS_plot <- data %>% ggplot(aes(JobSatisfaction))+geom_bar(aes(fill=Attrition)) + labs(title="JobSatisfaction vs Attrition")
JR_plot <-data %>% ggplot(aes(JobRole))+geom_bar(aes(fill=Attrition)) + labs(title="JobRole vs Attrition")
MS_plot <- data %>% ggplot(aes(MaritalStatus))+geom_bar(aes(fill=Attrition)) + labs(title="MaritalStatus vs Attrition")

#Chi square test to determine correlation
JL_chi <- chisq.test(data$JobLevel, data$Attrition)
O_chi <- chisq.test(data$OverTime, data$Attrition)
JI_chi <- chisq.test(data$JobInvolvement, data$Attrition)
JR_chi <- chisq.test(data$JobRole, data$Attrition)
MS_chi <- chisq.test(data$MaritalStatus, data$Attrition)

cat_var = c("JobLevel", "OverTime", "JobInvolvement", "JobRole", "MaritalStatus")
chi_p = c(JL_chi$p.value, O_chi$p.value, JI_chi$p.value, JR_chi$p.value, MS_chi$p.value)
df_chitest = data.frame(Variable=cat_var, Chisq.pvalue=chi_p)

grid.arrange(JL_plot, O_plot, JI_plot, JR_plot, MS_plot,tableGrob(arrange(df_chitest, Chisq.pvalue)))

#Doesn't appear
#data %>% group_by(PerformanceRating) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(RelationshipSatisfaction) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(Education) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(EducationField) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(Gender) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(BusinessTravel) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(Department) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(WorkLifeBalance) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#data %>% group_by(JobSatisfaction) %>% count(Attrition) %>% mutate(sum=sum(n)) %>% mutate(perc_by_cat=n/sum) %>% filter(Attrition=="Yes")
#Attrition model plots

```
  
##NaiveBayes
The NB model was created by adding a signficant variable one at a time.  
Ulitmately 7 variables were used to reach Sensitivity and Specificity >60%.  
These variables were:  
* OverTime  
* JobRole  
* JobInvolvement  
* MonthlyIncome  
* TotalWorkingYears  
* YearsInCurrentRole  
* JobLevel  
The Confusion matrix and plots are below.  
```{r Attrition model}

set.seed(6)
splitPerc = .75

trainindex = sample(seq(1,dim(data)[1],1), round(splitPerc*dim(data)[1]))

trainIndices = sample(seq(1,dim(data)[1],1),round(splitPerc * dim(data)[1]))
train = data[trainindex,]
test = data[-trainindex,]

m = naiveBayes(Attrition~OverTime+JobRole+JobInvolvement+MonthlyIncome+TotalWorkingYears+YearsInCurrentRole+JobLevel,data=train)
table(predict(m, newdata=test),test$Attrition)
CM = confusionMatrix(table(predict(m, newdata=test),test$Attrition))
CM

grid.arrange(O_plot, JR_plot,JI_plot, MI_plot, TWY_plot,YICR_plot,JL_plot)

comp_NB = naiveBayes(Attrition~OverTime+MonthlyIncome+JobRole+TotalWorkingYears+JobInvolvement+YearsInCurrentRole+JobLevel,data=data)

pred_att = data.frame(Attrition =predict(comp_NB, newdata=Comp_attr))
att_comp <-bind_cols(Comp_attr,pred_att)

#summary(att_comp)

#write.csv(att_comp,file="Case2PredictionsJohnson Attrition.csv")
```
  
## Salary EDA
4 Categorical variables were observed to have some type of relationship with MonthlyIncome.  
* JobLevel  
* Education  
* JobRole  
* EducationField  
The boxplots of these variables are below:  
```{r Salary EDA}

#Ordinal variable EDA
#data %>% select(Education, EnvironmentSatisfaction, JobInvolvement, JobSatisfaction, PerformanceRating, WorkLifeBalance, JobLevel, RelationshipSatisfaction, MonthlyIncome) %>% ggpairs(upper = list(continuous="smooth", combo="box", discrete = "facetbar"), lower=list(continuous="smooth", combo="box", discrete = "facetbar"))
# Correlation to JobLevel, maybe Education

JLS_plot <- data %>% ggplot(aes(JobLevel, MonthlyIncome))+geom_boxplot()+labs(title="Job Level vs Monthly Income")

ES_plot <- data %>% ggplot(aes(Education, MonthlyIncome))+geom_boxplot()+labs(title="Education vs Monthly Income")

#Nominal variable EDA
#data %>% select(Attrition, BusinessTravel, Department, EducationField, Gender, JobRole, MaritalStatus, OverTime, MonthlyIncome) %>% ggpairs(upper = list(continuous="smooth", combo="box", discrete = "facetbar"), lower=list(continuous="smooth", combo="box", discrete = "facetbar"))
# JobRole has interesting differences.  Maybe EducationField

JRS_plot <- data %>% ggplot(aes(JobRole,MonthlyIncome))+geom_boxplot()+labs(title="JobRole vs Monthly Income")

EFS_plot <- data %>% ggplot(aes(EducationField,MonthlyIncome))+geom_boxplot()+labs(title="EducationField vs Monthly Income")

grid.arrange(JLS_plot, ES_plot, JRS_plot, EFS_plot)
```
  
3 Continuous variables were observed to have some type of relationship with MonthlyIncome.  
* Age  
* TotalWorkingYears  
* Years at Company  
This was confirmed by a scatterplot for each.  
A correlation matrix was completed with additional variables to confirm as well.  
```{r Salary EDA Cont}
#Continous variable EDA
#data %>% select(Age,DailyRate, DistanceFromHome, HourlyRate, MonthlyRate, NumCompaniesWorked, PercentSalaryHike, MonthlyIncome) %>% ggpairs(upper = list(continuous="smooth", combo="box", discrete = "facetbar"), lower=list(continuous="smooth", combo="box", discrete = "facetbar"))
# Age has a positive correlation

AgS_plot <- data %>% ggplot(aes(Age,MonthlyIncome))+geom_point()+geom_smooth(method="lm")+labs(title = "Age vs Monthly Income")

#data %>% select(StockOptionLevel, TotalWorkingYears, TrainingTimesLastYear, YearsSinceLastPromotion, YearsInCurrentRole, YearsWithCurrManager, YearsAtCompany, MonthlyIncome) %>% ggpairs(upper = list(continuous="smooth", combo="box", discrete = "facetbar"), lower=list(continuous="smooth", combo="box", discrete = "facetbar"))
#TotalWorking Years and YearsatCompany have a strong correlation
#Years SinceLastPromotion, InCurrentRole and WithCurrManager have weak positive correlation

TWYS_plot <- data %>% ggplot(aes(TotalWorkingYears,MonthlyIncome))+geom_point()+geom_smooth(method="lm")+labs(title = "TotalWorkingYears vs Monthly Income")

YaCS_plot <- data %>% ggplot(aes(YearsAtCompany,MonthlyIncome))+geom_point()+geom_smooth(method="lm")+labs(title = "YearsAtCompany vs Monthly Income")

grid.arrange(AgS_plot,TWYS_plot, YaCS_plot)

corr_data <- data %>% select(Age, TotalWorkingYears, YearsAtCompany, YearsSinceLastPromotion, YearsInCurrentRole, YearsWithCurrManager, MonthlyIncome)

corr <- round(cor(corr_data),1)

ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of Continuous Variables", 
           ggtheme=theme_bw)
#TotalWorkingYears

```
## Salary Regression
The Regression model was build by adding one variable at a time.  
Two variables were used to achieve the RMSE < $3000.  
These variables were:  
* TotalWorkingYears  
* JobLevel  
The plots and regression results are below.  
```{r Salary Model}
Salary_train <- data %>% select(MonthlyIncome, TotalWorkingYears, JobLevel)

fit <- lm(MonthlyIncome~TotalWorkingYears+JobLevel, data=Salary_train)

summary(fit)
#confint(fit)

train(MonthlyIncome~TotalWorkingYears+JobLevel, method="lm",data=Salary_train, trControl = trainControl(method = "LOOCV"))

grid.arrange(TWYS_plot, JLS_plot)

pred_sal = data.frame(MonthlyIncome = predict(fit, newdata = Comp_sal))
sal_comp <-bind_cols(Comp_sal, pred_sal)

#summary(sal_comp)

#write.csv(sal_comp, file="Case2PredictionsJohnson Salary.csv")

```

##In Summary
Removed 3 variables from consideration (StandardHours, Over18, EmployeeCount)  
10 Variables considered for Attrition Prediction Model  
* 5 Categorical  
* 5 Continous  
Final Attrition Prediction Model had 7 Variables(4 Categorical/3Continuous)  
*Sensitivity = 88%   
*Specificity = 62%   
7 Variables considered for Salary Model  
* 4 Categorical Variables  
* 3 Continuous Variables  
Final Salary Model has 2 Variables (1 each Categorical/Continuous)  
* R2 = .93
* RMSE = $1262  
